---
title: "Lab 03 - Nobel laureates"
author: "George Nesbitt"
date: "01/28/2026"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1
There are 935 observations and 26 variables in the nobel dataset
```{r glimpse of nobel}
glimpse(nobel)
```

### Exercise 2
```{r new data frame}
nobel_living <- nobel %>%
  filter(is.na(died_date),!is.na(country),gender != "org")
```
```{r glimpse of new data frame}
glimpse(nobel_living)
```


### Exercise 3
```{r visualization 1}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```
```{r}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```
```{r}
ggplot(data=nobel_living_science, mapping=aes(y=category))+geom_bar()+facet_wrap(~country_us)
```

There were more nobel laurates that won their prizes in the US than in other countries. Specifically, economists were 
almost primarily in the United States. This supports the Buzzfeed headline. 

### Exercise 4
```{r new variable}
nobel_living <- nobel_living %>%
  mutate(
    born_country_us = if_else(born_country == "USA", "USA", "Other")
  )
```
```{r}
nobel_living%>%
count(born_country_us)
```
There are 105 living nobel laureates born in the United States.

...

### Exercise 5
```{r visualization 2}
ggplot(
  data = nobel_living,
  mapping = aes(x = country_us, fill = born_country_us)
) +
  geom_bar() +
  facet_wrap(~ category)

```
Based on this visualization, it appears that, compared to the data we found from ex 3, Buzzfeed's claim that of US-based Nobel laureates, many were born in other countries is true. It does appear that for the most part, there are more international-born Nobel laureates than US laureates. 
...

### Exercise 6
```{r new filter for non US born American laureates}
nobel_living %>%
  filter(country_us == "USA", born_country !="USA") %>%
  count(born_country) %>%
  arrange(desc(n))
```

The countries that are the most common are Germany and the UK. 
...
